\documentclass[solution, letterpaper]{cs20inclass}
\usepackage{enumerate}
\usepackage{tikz}
\usepackage{pgf}
\usepackage{tikz}
\usepackage{hyperref}
\begin{document}
\header{17}{Monday, March 7, 2016}

\noindent Author: Michelle Danoff, Tom Silver% \\

\paragraph*{Executive Summary}
\begin{enumerate}
\item A \textit{directed graph} $G$ consists of a set of \textit{vertices} $V$ and a set of \textit{edges} $E \subset V \times V$. $G$ is a binary relation on $V$, where $uGw$ iff there is an edge from $u$ to $w$.
\begin{itemize}
\item A directed edge $(u,v) \in E$ means that there is an edge pointing from vertex $u$ to vertex $v$ in the graph.
\item The \textit{in-degree} of a vertex is the number of edges pointing into it. The \textit{out-degree} of a vertex is the number of edges pointing out of it.
\end{itemize}
\item A \textit{walk} is a sequence of vertices in the directed graph obtained by ``walking'' along the edges from one vertex to the next in the sequence. A \textit{path} is a walk that does not visit the same vertex more than once.
\begin{itemize}
\item The shortest walk between two vertices is a path. 
\end{itemize}
\item The binary relations $G^*$ and $G^+$ are defined such that for $u, v \in V$:
\begin{itemize}
\item $u G^* v$ means that there is a walk of length $\ge 0$ in $G$ from $u$ to $v$
\item $u G^+ v$ means that there is a positive length walk in $G$ from $u$ to $v$
\end{itemize}
\item A \textit{cycle} is a walk from a vertex to itself, with no other repeated vertices in the walk. A single vertex is a length 0 cycle.
\begin{itemize}
\item A \textit{closed walk} is a walk that begins and ends at the same vertex. The shortest positive length closed walk through a vertex is a positive length cycle containing that vertex.
\end{itemize}
\item A \textit{directed acyclic graph} (DAG) contains no positive length cycles.

\end{enumerate}

\problem
Prove that any finite DAG has at least one vertex with in-degree 0.


\begin{solution}
  % Write your answer here.
\end{solution}

\problem
Consider the following set of prerequisites for CS classes (loosely adapted from the course catalog; not all of these are actually strict prerequisites): \\
Prerequisites: \\
CS 182: CS 51, CS 121 \\
CS 121: CS 20 \\
CS 124: CS 50, CS 51, CS 121, Stat 110 \\
CS 51: CS 50 \\
CS 61: CS 50 \\
CS 20, CS 50, \& Stat 110 have no prerequisites, but should be in the graph.

\subproblem Draw the directed graph representing these prerequisites, where the vertices represent classes and where an edge from $x$ to $y$ means that $x$ is a prerequisite for $y$.
\subproblem Is this graph a directed acyclic graph (DAG)? Why does it not make sense for a graph like this representing prerequisites to contain cycles?
\subproblem If you must take all of a class's prerequisite classes before taking the class, what is the fewest possible number of semesters it would take to complete all these courses? Assume you are able to handle an unlimited workload every semester.

\begin{solution}
  % Write your answer here.
\end{solution}

\problem Prove the triangle inequality, which states that for any $u, x, v \in V$
\[dist(u, v) \le dist(u, x) + dist(x, v)\]
where $dist(u,v)$ is the length of the shortest path between the vertices $u$ and $v$.

\begin{solution}
  % Write your answer here.
\end{solution}

\problem [BONUS] A topological ordering of a graph is an ordering of the nodes such that if there is an edge between vertexes a and b in the graph, then a is before b in the topological ordering. Prove that we can provide a topological ordering for any DAG. 

\begin{solution}
  % Write your answer here.
\end{solution}


\end{document}
